A Methodology and Implementation of Automated Emissions Harmonization for Use in Integrated Assessment Models

### 2017-06-06 intro.tex
addresed shinichiro's comments

Integrated Assessment Models (IAMs) are tools used to understand the complex
interactions between energy systems, economic systems, land use, the climate,
and air pollution, among other sectors . IAMs provide global projections of
systemic change by dividing the world into a number of representative regions,
the definition of which is distinct for each model. Results from IAMs are
integral in a number of international studies, which notably include projections
of climate and economic futures. Recently, flagship IAMs have developed
scenarios based on the Shared Socioeconomic Pathways (SSPs)
\cite{van_vuuren_energy_2017, kriegler_fossil-fueled_2017, calvin_ssp4:_2017,
  fujimori_ssp3:_2017, fricko_marker_2017} which model a variety of potential
global futures. The SSPs are used in analyses underpinning the Intergovernmental
Panel on Climate Change (IPCC) and the United Nations Framework Convention on
Climate Change (UNFCCC)'s Paris Agreement \TODO{cite}.

While IAMs are implemented in myriad ways, including simulation and
optimization, the core inputs, outputs, and purpose are similar across different
models. Modeling teams incorporate existing datasets of energy systems, world
development indicators, and emissions sources and concentrations, among other
data, in order to provide consistent historical trajectories of modeled
variables. The models then provide estimates of future trajectories of these
variables under various socioeconomic and technological assumptions as well as
proposed policy constraints, e.g., targets for future Greenhouse Gas (GHG)
emissions.

Future emissions pathways are of critical importance to both of global and local
governance concerns. For example, the United Nation's Sustainable Development
Goals (SDGs) \TODO{CITE} speak directly to the impacts of emissions in Goal 3
(Good Health and Well-Being) and Goal 13 (Climate Action). IAMs endogenously
calculate the trajectories many of the emissions species of import that impact
the progress towards these SDGs by modeling the individual technologies and
sectors that contribute towards the emissions of different air pollutants and
GHGs as well as various mitigation technologies (e.g., NOx and SOx scrubbers on
power plants).

@@

In order to incorporate the large variety of possible model results and given
relatively large increases in regional and sectoral dimensionality, an automated
process for determining harmonization methods is needed. The remainder of this
paper describes the methodology of the \code{aneris} code base, written in the
Python programming language, which implements such a system in Section
\ref{sec:meths}. The results of applying the automated harmonization mechanism
on two example IAM scenarios, one with emissions growth and another with
emissions mitigation, is presented in Section \ref{sec:results}. Finally, the
general effectiveness and potential future improvements on the automated
methodology is discussed in Section \ref{sec:future}.

>>>>

Integrated Assessment Models (IAMs) are tools used to understand the complex
interactions between energy systems, economic systems, land use, the climate,
and air pollution, among other sectors . IAMs provide global projections of
systemic change by dividing the world into a number of representative regions
(typically 10 to 30), the definition of which is distinct for each model
\cite{krey_global_2014}. Results from IAMs are integral in a number of
international studies, which notably include projections of climate and economic
futures. Recently, flagship IAMs have developed scenarios based on the Shared
Socioeconomic Pathways (SSPs) \cite{van_vuuren_energy_2017, fricko_marker_2017,
  fujimori_ssp3:_2017, calvin_ssp4:_2017, kriegler_fossil-fueled_2017} which
quantify a variety of potential global futures. The SSPs are primarily designed
to being used in analyses of broad climate research which includes earth system
model (ESM) simulations, climate impact, adaptation climate mitigation
studies. They consequently underpin the United Nations' Intergovernmental Panel
on Climate Change (IPCC) across working groups.

While IAMs are implemented in myriad ways, including simulation and
optimization, the core inputs, outputs, and purpose are similar across different
models. Modeling teams incorporate existing data sets of energy systems, land
use, economic, demographic and emissions sources and concentrations, among other
data, in order to provide consistent historical trajectories of modeled
variables. The models then provide estimates of future trajectories of these
variables under various socioeconomic and technological assumptions as well as
proposed policy constraints, e.g., targets for future Greenhouse Gas (GHG)
emissions.

The emissions trajectories calculated by IAMs are critical inputs for ongoing,
worldwide scientific community efforts in the Coupled Model Intercomparison
Project Phase 6 (CMIP6) \cite{eyring_overview_2016}, which is utilizing a number
of marker SSP scenarios developed by the IAM community
\cite{oneill_scenario_2016}. These trajectories are endogenously calculated by
modeling the individual technologies and sectors that contribute towards the
emissions of different air pollutants and GHGs as well as various mitigation
technologies (e.g., NOx and SOx scrubbers on power plants). However, the
historical emissions inventories used by different models can differ by large
amounts depending on the region, sector, and emissions species being modeled.

@@

In previous studies, e.g. the Representative Concentration Pathways (RCPs), IAM
results have been harmonized by sector and the 5 RCP global regions
\cite{vuuren_representative_2011}. Climate modeling efforts have continued to
progress, however, demanding increased spatial and sectoral resolution. In order
to address the growing dimensionality of model outputs while still providing a
consistent and scientifically rigorous harmonization procedure, an automated
process for determining harmonization methods is needed. The remainder of this
paper describes the methodology of the \code{aneris} code base
\cite{matthew_gidden_2017_802832}, written in the Python programming language,
which implements such a system in Section \ref{sec:meths} (documentation
available \href{http://mattgidden.com/aneris/}{online}). The results of applying
the automated harmonization mechanism on two example IAM scenarios, one with
emissions growth and another with emissions mitigation, is presented in Section
\ref{sec:results}. Finally, the general effectiveness and potential future
improvements on the automated methodology is discussed in Section
\ref{sec:future}.

### 2017-06-06 meth.tex
more meth revisions
Shinichiro_30May
addresed shinichiro's comments

IAM emission results are provided along a number of different dimensions,
including, temporal (normally half decade or decade), spatial (i.e., model
regions), gas species, and sectoral dimensions. Each individual temporal
trajectory, i.e., unique combinations of regions ($r$), species ($g$), and
sectors ($s$), must be harmonized to the initial time step. Given a model
trajectory, $m_{r, g, s}(t)$, historical trajectory, $h_{r, g, s}(t)$, and
initial time step $t_i$, a harmonized trajectory can be calculated.

@@

The full \code{aneris} workflow is comprised of a number of components shown
graphically in Figure \ref{fig:workflow}. Unharmonized model data and a
run-control configuration are read in via an Excel spreadsheet. Data is assumed
to be in the \textit{IAM Consortium} format, i.e., using \code{Model},
\code{Scenario}, \code{Region}, \code{Variable}, and \code{Unit} columns in
addition to columns representing each modeled time period. Users are able to
control the harmonization process via a number of options. First, users can
provide \textit{override} methods for any combination of region and variable
(i.e., sector and gas species). If override methods are provided, they are used
instead of the default methods as determined by the default method decision
tree. Furthermore, users can set the above-mentioned thresholds as well as the
LUC method used in the decision tree (see Figure
\ref{fig:fig:decision_tree}). Further detail of input parameters can be found at
\hyperlink{http://aneris.readthedocs.io/en/latest/}.

@@

Input data then undergoes a cleaning operation, which adds (null) model
trajectories that exist in the historical dataset but are not provided by the
model input and detects any issues that would cause the harmonization process to
fail. The methods used to harmonize the data are then determined and the
harmonization process is executed. Upon completion of the harmonization process,
aggregation of common analysis regions is performed (e.g., the 5 SRES
regions). Finally, any exogenous trajectories the user provides are
added. Exogenous trajectories are normally provided for unmodeled gases with
well-accepted scenario trajectories, e.g., chlorofluorocarbons provided by WMO
\TODO{CITE}. Upon completion, the harmonized trajectories and meta data
regarding the harmonization process are returned. A description of all returned
meta data is provided in Table \ref{tab:metadata}.

>>>>

IAM emission results are provided along a number of different dimensions,
including, temporal (normally half decade or decade), spatial (i.e., model
regions), gas species, and sectoral dimensions. Each individual temporal
trajectory, i.e., unique combinations of regions ($r$), species ($g$), and
sectors ($s$), must be harmonized to the initial time step. Given a model
trajectory, $m_{r, g, s}(t)$, historical trajectory, $h_{r, g, s}(t)$, and
initial time step $t_i$, a harmonized trajectory can be calculated. The
``goodness of fit'' of a trajectory depends on a number of factors, most notably
the faithful representation of the original, unharmonized, trajectory. However,
other important factors must be considered, such the representation of negative
trajectories (i.e., if a trajectory becomes negative, both the timing and total
magnitude should be as close as possible) which are of critical importance for
cumulative CO2 calculations.

@@

The full \code{aneris} workflow is comprised of a number of components shown
graphically in Figure \ref{fig:workflow}. Unharmonized model data and a
run-control configuration are read in via an Excel spreadsheet. Data is assumed
to be in the \textit{IAM Consortium} format, i.e., using \code{Model},
\code{Scenario}, \code{Region}, \code{Variable}, and \code{Unit} columns in
addition to columns representing each modeled time period. Users are able to
control the harmonization process via a number of options. First, users can
provide \textit{override} methods for any combination of region and variable
(i.e., sector and gas species). If override methods are provided, they are used
instead of the default methods as determined by the default method decision
tree. Furthermore, users can set the above-mentioned thresholds as well as the
LUC method used in the decision tree (see Figure
\ref{fig:decision_tree}). Further detail of input parameters can be found at
\hyperlink{http://aneris.readthedocs.io/en/latest/}{http://aneris.readthedocs.io/en/latest/}.

@@

Input data then undergoes a cleaning operation, which adds (null) model
trajectories that exist in the historical dataset but are not provided by the
model input and detects any issues that would cause the harmonization process to
fail. The methods used to harmonize the data are then determined and the
harmonization process is executed. Upon completion of the harmonization process,
aggregation of common analysis regions is performed (e.g., the 5 SRES
regions). Finally, any exogenous trajectories the user provides are
added. Exogenous trajectories are normally provided for unmodeled gases with
well-accepted scenario trajectories, e.g., chlorofluorocarbons provided by WMO
\cite{wmo2014}. Upon completion, the harmonized trajectories and meta data
regarding the harmonization process are returned. A description of all returned
meta data is provided in Table \ref{tab:metadata}.
